GMU MS SWE Handbook\\{\large A Handbook For the Masters of Software Engineering program at George Mason University

### 2024-08-07 nnv.tex
create NNV book

\paragraph{School} Graduating from top universities \emph{that we recognize} helps. For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is a plus.
However, if we do not know much about schools in your country, then we are uncertain about the quality of your school and likely treat your school as \emph{``unknown foreign''}, which can be a minus point.

@@

\paragraph{Grades} Having a good GPA is good. However if your school is not well-known, having top grades or rankings
usually will not help because we cannot evaluate them.  This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by adcom).

As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.
Note that while having good grades might not help much,
having bad grades (e.g., < 3.0) will be \red{red flag} (unless your LORs or
statement give a proper explanation). This is especially true if you
have bad grades in relevant courses, e.g., Math and CS.

@@

>>>>

\subsection{School} Graduating from top universities \emph{that we recognize} helps. For example, if you are an international student and your school is well-known, then it is \emph{``top foreign''}, which is a plus.
However, if we do not know much about schools in your country, then we are uncertain about the quality of your school and likely treat your school as \emph{``unknown foreign''}, which can be a minus point.

@@

\subsection{Grades} Grades generally do not matter much for CS PhD admission. PhD in CS is a research degree and doing well in undergrad courses does not necessarily mean you can do research (other factors such as research experiences, pubs, LoRs are more important).

Nonetheless, if you are from a well-known school, having good grades do help (not a lot though), e.g., adcom members often note details such as "good GPA from well-known school (\autoref{sec:ievaluate}). However if your school is not well-known, having top grades or rankings usually will not help because we cannot evaluate them (e.g., we don't know how hard it is to get a 4.0 or A's at your school). This can be an issue for students in many top international universities where the competition is so high that very good students can still have low rankings from these schools (and be overlooked by adcom).

As with school reputation, you and your LoR writers can mention the grading system of your university if you think that is helpful for adcom to evaluate.

@@

\paragraph{Bad Grades} 
While having good GPA might not help much (again, because of other much more important factors),
having bad GPA can be a \red{red flag}. In fact, many universities have a minimum GPA requirement (e.g., > 3.0) and will automatically reject applications with lower GPAs.  
If you have bad grades, you should explain them in your SOP or have your LOR writers explain them in their letters.

Moreover, having bad grades in relevant courses, e.g., Math and CS, can be a \red{red flag}.
Adcom members often scan through transcripts looking for C and lower grades in Math and CS courses and might raise concerns if they see several of them.
Note that bad grades in non-relevant courses, e.g., e.g., about politics or history, are not as concerning.

A common question is whether you should be concern and explain bad grades in relevant courses in your SOP.  If you have just a few, then they do not matter much, so don't spend much time explaining them (let's face it, many of us, including faculty, have bad grades in some courses, including relevant ones). But if you have many bad grades for an entire semester or year due to some specific reasons, then you may be you should explain them in your SOP.
